\documentclass[12pt]{article}

\usepackage{graphicx}
\usepackage{paralist}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{hhline}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{url}

\oddsidemargin -10mm
\evensidemargin -10mm
\textwidth 160mm
\textheight 200mm
\renewcommand\baselinestretch{1.0}

\pagestyle {plain}
\pagenumbering{arabic}

\newcounter{stepnum}

%% Comments

\usepackage{color}

\newif\ifcomments\commentstrue

\ifcomments
\newcommand{\authornote}[3]{\textcolor{#1}{[#3 ---#2]}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
\else
\newcommand{\authornote}[3]{}
\newcommand{\todo}[1]{}
\fi

\newcommand{\wss}[1]{\authornote{blue}{SS}{#1}}

\title{Assignment 3, Part 1, Specification}
\author{SFWR ENG 2AA4\\Jay Mody - 400195508 - modyj}
\date{\today}

\begin {document}

\maketitle
This Module Interface Specification (MIS) document contains modules, types and
methods for implementing a generic 2D sequence that is instantiated for both
land use planning and for a Discrete Elevation Model (DEM).

In applying the specification, there may be cases that involve undefinedness.
We will interpret undefinedness following~\cite{Farmer2004}:

If $p: \alpha_1 \times .... \times \alpha_n \rightarrow \mathbb{B}$ and any of
$a_1, ..., a_n$ is undefined, then $p(a_1, ..., a_n)$ is False.  For instance,
if $p(x) = 1/x < 1$, then $p(0) = \text{False}$.  In the language of our
specification, if evaluating an expression generates an exception, then the
value of the expression is undefined.

\wss{The parts that you need to fill in are marked by comments, like this one.
  In several of the modules local functions are specified.  You can use these
  local functions to complete the missing specifications.}

\wss{As you edit the tex source, please leave the \texttt{wss} comments in the
  file.  Put your answer \textbf{after} the comment.  This will make grading
  easier.}

\bibliographystyle{plain}
\bibliography{SmithCollectedRefs}

\newpage

\section* {Land Use Type Module}

\subsection*{Module}

LanduseT

\subsection* {Uses}

N/A

\subsection* {Syntax}

\subsubsection* {Exported Constants}

None

\subsubsection* {Exported Types}

Landtypes = \{R, T, A, C\}\\

\noindent \textit{//R stands for Recreational, T for Transport, A for Agricultural, C for
  Commercial}

\subsubsection* {Exported Access Programs}

\begin{tabular}{| l | l | l | p{5cm} |}
\hline
\textbf{Routine name} & \textbf{In} & \textbf{Out} & \textbf{Exceptions}\\
\hline
new LanduseT & Landtypes & LanduseT & ~\\
\hline
\end{tabular}

\subsection* {Semantics}

\subsubsection* {State Variables}

landuse: Landtypes

\subsubsection* {State Invariant}

None

\subsubsection* {Access Routine Semantics}

\noindent new LandUseT($t$):
\begin{itemize}
\item transition: $\mathit{landuse} := t$
\item output: $out := \mbox{self}$
\item exception: none
\end{itemize}

\subsubsection* {Considerations}

When implementing in Java, use enums (as shown in Tutorial 06 for ElementT).

\newpage

\section* {Point ADT Module}

\subsection*{Template Module inherits Equality(PointT)}

PointT

\subsection* {Uses}

N/A

\subsection* {Syntax}

\subsubsection* {Exported Types}

\wss{What should be written here?} PointT = ?

\subsubsection* {Exported Access Programs}

\begin{tabular}{| l | l | l | l |}
\hline
\textbf{Routine name} & \textbf{In} & \textbf{Out} & \textbf{Exceptions}\\
\hline
PointT & $\mathbb{Z}$, $\mathbb{Z}$ & PointT & \\
\hline
row & ~ & $\mathbb{Z}$ & ~\\
\hline
col & ~ & $\mathbb{Z}$ & ~\\
\hline
translate & $\mathbb{Z}$, $\mathbb{Z}$ & PointT & ~\\
\hline
\end{tabular}

\subsection* {Semantics}

\subsubsection* {State Variables}

$r$: \wss{What is the type of the state variables?} $\mathbb{Z}$\\
$c$: \wss{What is the type of the state variables?} $\mathbb{Z}$

\subsubsection* {State Invariant}

None

\subsubsection* {Assumptions}

The constructor PointT is called for each object instance before any other
access routine is called for that object.  The constructor cannot be called on
an existing object.

\subsubsection* {Access Routine Semantics}

PointT($row, col$):
\begin{itemize}
\item transition: \wss{What should the state transition be for the constructor?}\\$r, c := row, col$
\item output: $out := \mathit{self}$
\item exception: None
\end{itemize}

\noindent row():
\begin{itemize}
\item output: $out := r$
\item exception: None
\end{itemize}

\noindent col():
\begin{itemize}
\item \wss{What should go here?} output: $out := c$
\item exception: None
\end{itemize}

\noindent translate($\Delta r$, $\Delta c$):
\begin{itemize}
\item \wss{What should go here?} output: $out := \mbox{PointT}(r + \Delta r, c + \Delta c)$
\item exception: \wss{What should go here?} None
\end{itemize}

\newpage

\section* {Generic Seq2D Module}

\subsection* {Generic Template Module}

Seq2D(T)

\subsection* {Uses}

PointT

\subsection* {Syntax}

\subsubsection* {Exported Types}

Seq2D(T) = ?

\subsubsection* {Exported Constants}

None

\subsubsection* {Exported Access Programs}

\begin{tabular}{| l | l | l | p{6cm} |}
\hline
\textbf{Routine name} & \textbf{In} & \textbf{Out} & \textbf{Exceptions}\\
\hline
Seq2D & seq of (seq of T), $\mathbb{R}$ & Seq2D & IllegalArgumentException\\
\hline
set & PointT, T & ~ & IndexOutOfBoundsException\\
\hline
get & PointT & T & IndexOutOfBoundsException\\
\hline
getNumRow & ~ & $\mathbb{N}$ & \\
\hline
getNumCol & ~ & $\mathbb{N}$ & \\
\hline
getScale & ~ & $\mathbb{R}$ & \\
\hline
count & T & $\mathbb{N}$ & \\
\hline
countRow & T, $\mathbb{N}$ & $\mathbb{N}$ & \\
\hline
area & T & $\mathbb{R}$ & \\
\hline
\end{tabular}

\subsection* {Semantics}

\subsubsection* {State Variables}

$s$: seq of (seq of T)\\
scale: $\mathbb{R}$\\
nRow: $\mathbb{N}$\\
nCol: $\mathbb{N}$

\subsubsection* {State Invariant}

None

\subsubsection* {Assumptions}

\begin{itemize}
\item The Seq2D(T) constructor is called for each object instance before any
other access routine is called for that object.  The constructor can only be
called once.
\item Assume that the input to the constructor is a sequence of rows, where each
  row is a sequence of elements of type T.  The number of columns (number of
  elements) in each row is assumed to be equal. That is each row
  of the grid has the same number of entries.  $s[i][j]$ means the ith row and
  the jth column.  The 0th row is at the top of the grid and the 0th column
  is at the leftmost side of the grid.
\end{itemize}

\subsubsection* {Access Routine Semantics}

Seq2D($S$, scl):
\begin{itemize}
\item transition: \wss{Fill in the transition.}\\
$s, scale, nRow, nCol := S, scl, |S|, |S[0]|$
\item output: $\mathit{out} := \mathit{self}$
\item exception: \wss{Fill in the exception.  One should be generated if the
    scale is less than zero, or the input sequence is empty, or the number of
    columns is zero in the first row, or the number of columns in any row is
    different from the number of columns in the first row.}\\ \\
\begin{tabular}{|p{6cm}|l|}
\hhline{~|-|}
\multicolumn{1}{r|}{} & \multicolumn{1}{l|}{$exc :=$}\\
\hhline{|-|-|}
$scl \leq 0$ & IllegalArgumentException \\
\hhline{|-|-|}
$|S| = 0$ & IllegalArgumentException \\
\hhline{|-|-|}
$|S[0]|=0$& IllegalArgumentException \\
\hhline{|-|-|}
$\exists\ \mbox{row} \in S \,.\, \lnot(|\mbox{row}| = |S[0]|)$ & IllegalArgumentException \\
\hhline{|-|-|}
\end{tabular}


\end{itemize}

\noindent set($p, v$):
\begin{itemize}
\item transition: \wss{?} $s[p.row()][p.col()] := v$
\item exception: \wss{Generate an exception if the point lies outside of the
    map.} \\
$exc := \lnot (\mbox{validPoint(p)}) \Rightarrow \mbox{IndexOutOfBoundsException}$
\end{itemize}

\noindent get($p$):
\begin{itemize}
\item output: \wss{?} $out := s[p.row()][p.col()]$
\item exception: \wss{Generate an exception if the point lies outside of the
    map.} \\
$exc := \lnot (\mbox{validPoint(p)}) \Rightarrow \mbox{IndexOutOfBoundsException}$
\end{itemize}

\noindent getNumRow():
\begin{itemize}
\item output: $out := \mbox{nRow}$
\item exception: None
\end{itemize}

\noindent getNumCol():
\begin{itemize}
\item output: $out := \mbox{nCol}$
\item exception: None
\end{itemize}

\noindent getScale():
\begin{itemize}
\item output: $out := \mbox{scale}$
\item exception: None
\end{itemize}

\noindent count($t$: T):
\begin{itemize}
\item output: \wss{Count the number of times the value $t$ occurs in the 2D
    sequence.} \\
$out := +(i,j : \mathbb{N} | \mbox{validRow(i)} \land \mbox{validCol(j)} \land s[i][j] = t : 1)$
\item exception: None
\end{itemize}

\noindent countRow($t$: T, $i: \mathbb{N}$):
\begin{itemize}
\item output: \wss{Count the number of times the value $t$ occurs in row
    $i$.} \\
$out := +(j : \mathbb{N} | \mbox{validCol(j)} \land s[i][j] = t : 1)$
\item exception: \wss{Generate an exception if the index is not a valid
    row.} \\
$exc := \lnot (\mbox{validRow(i)}) \Rightarrow \mbox{IndexOutOfBoundsException}$
\end{itemize}

\noindent area($t$: T):
\begin{itemize}
\item output: \wss{Return the total area in the grid taken up by cell value $t$.
    The length of each side of each cell in the grid is
    scale.} \\
$out := +(i,j : \mathbb{N} | \mbox{validRow(i)} \land \mbox{validCol(j)} \land s[i][j] = t : \mbox{scale}^2)$
\item exception: None
\end{itemize}

\subsection*{Local Functions}

\noindent validRow: $\mathbb{N} \rightarrow \mathbb{B}$\\
\noindent \wss{returns true if the given natural number is a valid row
  number.}\\
$\mbox{validRow(n)} \equiv 0 \leq n \leq (\mbox{nRow} - 1)$\\

\noindent validCol: $\mathbb{N} \rightarrow \mathbb{B}$\\
\noindent \wss{returns true if the given natural number is a valid column
  number.}\\
$\mbox{validCol(n)} \equiv 0 \leq n \leq (\mbox{nCol} - 1)$\\

\noindent validPoint: $\mbox{PointT} \rightarrow \mathbb{B}$\\
\noindent \wss{Returns true if the given point lies within the boundaries of the
  map.}\\
$\mbox{validPoint(p)} \equiv \mbox{validRow(p.row())} \land \mbox{validCol(p.col())}$\\

\newpage

\section* {LanduseMap Module}

\subsection* {Template Module}

\wss{Instantiate the generic ADT Seq2D(T) with the type LanduseT}\\
LanduseMapT is Seq2D(LanduseT)

\newpage

\section* {DEM Module}

\subsection* {Template Module}

DemT is Seq2D($\mathbb{Z}$)

\subsection* {Syntax}

\subsubsection* {Exported Access Programs}

\begin{tabular}{| l | l | l | p{6cm} |}
\hline
\textbf{Routine name} & \textbf{In} & \textbf{Out} & \textbf{Exceptions}\\
\hline
total & & $\mathbb{Z}$ & \\
\hline
max &  & $\mathbb{Z}$ & \\
\hline
ascendingRows & & $\mathbb{B}$ & \\
\hline
\end{tabular}

\subsection* {Semantics}

\subsubsection* {Access Routine Semantics}

\noindent total():
\begin{itemize}
\item output: \wss{Total of all the values in all of the cells.}\\
$out := +(i,j : \mathbb{N} | \mbox{validRow(i)} \land \mbox{validCol(j)} : s[i][j])$
\item exception: None
\end{itemize}

\noindent max():
\begin{itemize}
\item output: \wss{Find the maximum value in the 2d grid of integers}\\
$\exists \ x,y : \mathbb{Z} \,.\, (\mbox{validRow}(x) \land \mbox{validCol}(y) \land \\ (\forall i,j : \mathbb{Z} \,.\, \mbox{validRow}(i) \land \mbox{validCol}(j) \land s[x][y] \geq s[i][j])) \Rightarrow out := s[x][y])$
\item exception: None
\end{itemize}

\noindent ascendingRows():
\begin{itemize}
\item output: \wss{Returns True if the sum of all values in each row increases
    as the row number increases, otherwise, returns False.}\\
$\forall \ i, j : \mathbb{Z} \,.\, \mbox{validRow}(i) \land \mbox{validRow}(i-1) \land \mbox{validCol}(j) \land +(s[i][j] - s[i-1][j]) > 0$
\item exception: None
\end{itemize}

\subsection*{Local Functions}

\noindent validRow: $\mathbb{N} \rightarrow \mathbb{B}$\\
\noindent \wss{returns true if the given natural number is a valid row
  number.}\\
$\mbox{validRow(n)} \equiv 0 \leq n \leq (\mbox{nRow} - 1)$\\

\noindent validCol: $\mathbb{N} \rightarrow \mathbb{B}$\\
\noindent \wss{returns true if the given natural number is a valid column
  number.}\\
$\mbox{validCol(n)} \equiv 0 \leq n \leq (\mbox{nCol} - 1)$\\

\newpage

\section*{Critique of Design}

\wss{Write a critique of the interface for the modules in this project.  Is there
anything missing?  Is there anything you would consider changing?  Why?  One
thing you could discuss is that the Java implementation, following the notes
given in the assignment description, will expose the use of ArrayList for Seq2D.
 How might you change this?  There are repeated local functions in two modules.
What could you do about this?}\\

The most notable mistake in the design specification is that it's missing 2 modules, the Equality module and the Landtypes module.
Equality is inherited by PointT and Landtypes is used by LanduseT, but neither are ever defined anywhere in the MIS. To make matter worse,
LanduseT is redundant. It's sole purpose is to store a Landtype variable, adding no additional functionality.
To remove this redundancy, it would be advantagous to use Landtype directly, removinng an unecessary layer of abstraction.
Additionally, there is a redundancy with validRow/validCol, as they are duplicated in Seq2D and DemT as local functions.
A better design would be to include validRow, validCol, and validPoint all as part of the interface for
Seq2D, so that any modules that inherit it will also have access to those functions.
Furthermore, the use of an ArrayList in the implementation for Seq2D as stated in the assignment description,
may not be the optimal option for the task. Seq2D is analagous to a matrix, with an equivalent number of
elements in each row. To make things simpler, we can make the use of a 2 dimensional array, rather than an ArrayList of ArrayLists.
\\\\
In addition to your critique, please address the following questions:

\begin{enumerate}
\item The original version of the assignment had an Equality interface defined
  as for A2, but this idea was dropped.  In the original version Seq2D inherited
  the Equality interface.  Although this works in Java with the LanduseMapT, it is
  problematic for DemT.  Why is it problematic?  (Hint: DEMT is instantiated
  with the Java type Integer.)\\

  Classes in Java (like Integer, ArrayList, etc ...) automatically inherit an equivalence method.
  By default, this method ".equals()", compares object ids, however, in many cases this method is overriden to test for equivalence for
  whatever data it is representing. Since LanduseT is a class we defined, this is no problem if we define our own equals method,
  but this may be problematic for DemT, as Integer already has it's own defined equals method.

\item Although Java has several interfaces as part of the standard language,
  such as the Comparable interface, there is no Equality interface.  Instead
  equals is provided through inheritance from Object.  Why do you think the
  Java language designers decided to use inheritance for equality, instead of
  providing an interface?

  One reason why the Java designers may have decided to use inheritance for equality is that
  it allows them to define the default equivalence behaviour of an object by if their id's are equivalent (aka the same object in memory).
  Another, more important reason, is that it reduces ambiguity and redundancy.
  Many objects may have an equality property, but not a less than or greater than property. To make
  things less ambigious, it's easier to have equality inherited and overridden than via some workaround
  with compareTo (or as a seperate interface). As for redundancy, equality is a very common property (like toString),
  so it's easier to have it as it's own inhereited method rather than as an interface.

\item The qualities of good module interface push the design of the interface in
  different directions. Why is it rarely possible to achieve a module interface
  that simultaneously is essential, minimal and general?

  To create the most general interface, you often really need to think ahead in the future
  and realize the most modular and encaplsulating form of the module. Often this involves cutting
  down many functions that may have been essential to the initial design, but not to it's
  more general version. Or, on the other hand, you need to define added functionality to
  encaplsulate a wider range of modules that may implement an interface. In this sense, designing
  interfaces that are general, essential, and minimal proves to be a difficult task. You may
  have to redesign you're interface, add new modules/interfaces, and go through many iterations
  before you can come to a healthy balance of all three.

\end{enumerate}

\end {document}
